---
title: "Hw10"
output: word_document
---

Question 14.1
Read breast cancer dataset:
```{r}
set.seed(42)
rm(list = ls())
bc <- read.table("breast-cancer-wisconsin.data.txt", stringsAsFactors = FALSE, sep=",")
summary(bc)
```
We can see that only V7 has missing data.
Find all indices and rows of the missing data "?" in column V7:
```{r}
indices = which(bc$V7 == "?") #indices
indices
bc[indices,] #look at the rows of missing data
```
Now we need to make sure the percentage of missing data is less than 5%
```{r}
length(indices)/nrow(bc)
```
There is only about 2.3% so we can use imputation.

To find out if the missing rows are biased in any way with respect to the response:
```{r}
bc_clean <- bc[-indices,] #data without missing values
bc_missing <- bc[indices,] #data with missing values
#table showing response variable for original:
prop.table(table(bc$V11)) 
```
```{r}
#table showing response variable for w/o missing data
prop.table(table(bc_clean$V11)) 
```
```{r}
#table showing response variable for with missing data
prop.table(table(bc_missing$V11)) 
```
Their ratios are:
```{r}
0.6552217/0.3447783 #original
0.6500732/0.3499268 #w/o missing
0.875/0.125 #with missing
```
Based on only simple calculation, benign(2) response tends to have more missing values. But it is only an observation that might not even be true or affect our analysis.

Q 14.1.1 Mean/Mode Imputation
impute data with mean:
```{r}
mean <- mean(as.numeric(bc[-indices,"V7"]))
mean #mean of rest of data except for missing
bc_impute_mean <- bc
bc_impute_mean[indices,]$V7 <- mean #convert missing data to mean
bc_impute_mean$V7 <- as.integer(bc_impute_mean$V7)
```
So bc_impute_mean is the data with imputed mean.

impute with mode:
```{r}
#the function to get mode:
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
mode <- as.numeric(Mode(bc[-indices,"V7"])) #get mode of rest of data except for missing
mode
bc_impute_mode <- bc
bc_impute_mode[indices,]$V7 <- mode
bc_impute_mode$V7 <- as.integer(bc_impute_mode$V7)
```
So bc_impute_mode is the data with imputed mode.

Q 14.1.2 Regression Imputation
build a regression model using the data without missing values:
While building regression model, we need to take out column V11, which is the response we are going to predict later, because we need to avoid leaking information.
```{r}
bc_regression <- bc[-indices,2:10] #bc data without id and response
bc_regression$V7 <- as.integer(bc_regression$V7)
model_regression <- lm(V7~V2+V3+V4+V5+V6+V8+V9+V10, data=bc_regression) #build a model on V7 with other predictors
summary(model_regression)
```
Thr breast cancer lump is a very complicated biological situation. Since the scope is out of my knowledge, I will only use the above regression instead of trying to "predict"" V7 using the "best" predictors.

Get predictions for missing data and impute the new data:
```{r}
predict_V7 <- predict(model_regression, newdata=bc[indices,]) #prediction for V7
bc_regression <- bc
#round V7 column to integers between 1-10:
bc_regression[indices,]$V7 <- round(predict_V7) #impute V7
bc_regression$V7 <- as.numeric(bc_regression$V7)
bc_regression$V7 <- as.integer(bc_regression$V7)
```
Next, make sure all V7 data are between 1-10:
```{r}
bc_regression$V7[bc_regression$V7 > 10] <- 10 #anything above 10 shoube 10
bc_regression$V7[bc_regression$V7 < 1] <- 1 #anything below 1 should be 1
```
So bc_regression is the data with imputed prediction from regression.

Q 14.1.3 Regression with Perturbation Imputation
Instead of static regression prediction, now we can add some"random noise"" to your predicted regression value. By using R's built-in rnorm function, we can sample a randon number from the normal distribution of our predicted data.
```{r}
set.seed(42)
#mean centered at predicted V7, with standard deviation of predicted V7
predict_perturb <- rnorm(nrow(bc[indices,]), predict_V7, sd(predict_V7))
predict_perturb
```
See the differences:
```{r}
predict_V7
```
Now we can impute data back and create a new data:
```{r}
bc_perturb <- bc
bc_perturb[indices,]$V7 <- round(predict_perturb) #impute the perturbation data
bc_perturb$V7 <- as.numeric(bc_perturb$V7)
bc_perturb$V7 <- as.integer(bc_perturb$V7)
#make sure imputed value are within 1-10:
bc_perturb$V7[bc_perturb$V7 > 10] <- 10
bc_perturb$V7[bc_perturb$V7 < 1] <- 1
```
So bc_perturb is the data with imputed prediction from regression with perturbation.

Q14.1.4 Compare the results and quality of classification model (using KNN)
(1)compare question 1,2,3
i.With data of imputed mean:
```{r}
set.seed(42)
library(kknn)
model_mean <- cv.kknn(V11~V2+V3+V4+V5+V6+V7+V8+V9+V10, bc_impute_mean, kcv=5, k=2) #train a cross-validation 5-fold kknn model with data of imputed mean
predicted <- as.integer(model_mean[[1]][,2] + 0.5) # round off to 2 or 4
predicted
accuracy <- sum(predicted == bc_impute_mean$V11)/nrow(bc_impute_mean)
accuracy #accuracy
```
The model successfully classified 94.85% of data.

ii.With data of imputed mode:
```{r}
set.seed(42)
model_mode <- cv.kknn(V11~V2+V3+V4+V5+V6+V7+V8+V9+V10, bc_impute_mode, kcv=5, k=2)
predicted <- as.integer(model_mode[[1]][,2] + 0.5) # round off to 2 or 4
accuracy <- sum(predicted == bc_impute_mode$V11)/nrow(bc_impute_mode)
accuracy
```
The model successfully classified 94.99% of data.

iii.With data of imputed regression prediction:
```{r}
set.seed(42)
model_re <- cv.kknn(V11~V2+V3+V4+V5+V6+V7+V8+V9+V10, bc_regression, kcv=5, k=2)
predicted <- as.integer(model_re[[1]][,2] + 0.5) # round off to 2 or 4
accuracy <- sum(predicted == bc_regression$V11)/nrow(bc_regression)
accuracy
```
The model successfully classified 94.56% of data.

iv. With data of imputed regression with perturbation:
```{r}
set.seed(42)
model_perturb <- cv.kknn(V11~V2+V3+V4+V5+V6+V7+V8+V9+V10, bc_perturb, kcv=5, k=2)
predicted <- as.integer(model_perturb[[1]][,2] + 0.5) # round off to 2 or 4
accuracy <- sum(predicted == bc_perturb$V11)/nrow(bc_perturb)
accuracy
```
The model successfully classified 94.42% of data.

(2) the data that remains after data points with missing values are removed:
```{r}
bc_removed <- bc[-indices,] #remove entire rows with missing data
bc_removed$V7 <- as.numeric(bc_removed$V7)
set.seed(42)
model_remove <- cv.kknn(V11~V2+V3+V4+V5+V6+V7+V8+V9+V10, bc_removed, kcv=5, k=2)
predicted <- as.integer(model_remove[[1]][,2] + 0.5) # round off to 2 or 4
accuracy <- sum(predicted == bc_removed$V11)/nrow(bc_removed)
accuracy
```
The model successfully classified 95.46% of data.

In summary, the data without missing values seem to perform slightly better than anyone else.

Question 15.1
